---
title: "CLG authorship analytics - user guide (part 2)"
author: "Erwan Moreau"
date: "1/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview

This is the second part of the user guide for "CLG authorship analytics". The first part is [here](user-guide-part1.html).

## Rmd

This document was generated from an [R Markdown](https://rmarkdown.rstudio.com/) source file. The source file is provided in the repository, it can be used to reproduce these experiments. It can be executed through the RStudio interface ("knit" button) or as follows:

```
rmarkdown::render('user-guide-part2.Rmd')
```

- Naturally the document can also be read as a regular documentation.
- **Important:** this Rmd document includes [bash chunks](https://bookdown.org/yihui/rmarkdown/language-engines.html). In order for these to work, the environment must have been configured as explained in [Requirements](#requirements) below.

### Options

The `.Rmd` source document can be configured by modifying the following lines:

```{r myoptions}
packages.path <- 'packages'
Sys.setenv(CLG_AUTHORSHIP_PACKAGES_PATH = packages.path)
output.dir <- '/tmp/user-guide-part2.output'
Sys.setenv(OUTPUT_DIR = output.dir)
delete.previous.output <- TRUE
# normally not needed with regular setup:
Sys.setenv(ADD_TO_PERL5LIB = '/home/moreaue/perl5/lib/perl5')
Sys.setenv(NCORES = 20)
Sys.setenv(TASKS_DIR= '/tmp/clg-authorship.tasks')
```

* Note: `ncores` should be set to represent the number of cores available on the machine.
* The role of `TASKS_DIR` is described below in the "Parallel/distributed processing" section.

In order to manually execute some of the commands below, it is recommended to assign the appropriate value to the above environment variables for the whole session, for example:

```{bash init0}
export CLG_AUTHORSHIP_PACKAGES_PATH=packages
export OUTPUT_DIR=user-guide-part1.output
```

### Initialization

```{bash init1,echo=delete.previous.output,eval=delete.previous.output}
rm -rf "$OUTPUT_DIR" "TASKS_DIR"
```

```{bash init2}
mkdir "$OUTPUT_DIR"
```


# Requirements

The software and data requirements are the same as described in [user guide part 1](user-guide-part1.html#requirements), except that **this part requires a lot more computational power**. It is strongly recommended to run these experiments on a machine with many cores and a large amount of RAM available.



